### 2018-05-08 doc/paper.tex
Did a "final" pass and edit.

As in~\eqref{de-casteljau-error}, the accuracy of the compensated
result~\eqref{de-casteljau-2-error} may be arbitrarily bad for ill-conditioned
polynomial evaluations. For example, as the condition number grows in
Figure~\ref{fig:jlcs-10}, some points have relative error exactly equal to
\(1\); this indicates that \(\mathtt{CompDeCasteljau}(p, s) = 0\), which is
a complete failure to evaluate the order of magnitude of \(p(s)\). For
root-finding problems \(\mathtt{CompDeCasteljau}(p, s) = 0\) when
\(p(s) \neq 0\) can cause premature convergence and incorrect results.
We describe how to defer rounding into progressively
smaller error groups and improve the accuracy of the computed result by a
factor of \(\mach\) for every error group added. So we derive
\texttt{CompDeCasteljauK}, a \(K\)-fold compensated de Casteljau algorithm
that satisfies the following a prior bound for any arbitrary integer \(K\):
\begin{equation}
  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \bigO{\mach^K}.
\end{equation}
This means that the computed value with \texttt{CompDeCasteljauK} is now
as accurate as the result of the de Casteljau algorithm performed in
\(K\) times the working precision with a final rounding back to the
working precision.

@@

Throughout this paper we assume that the computation in floating point
arithmetic obeys the model
\begin{equation}
  a \star b = \fl{a \circ b} = (a \circ b)(1 + \delta_1) =
  (a \circ b) / (1 + \delta_2)
\end{equation}
where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
\in \left\{+, -, \times, \div\right\}\) and \(\left|\delta_1\right|,
\left|\delta_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
and \(\star\) represents the floating point computation, e.g.
\(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
\(\mach = 2^{-53}\).) We denote the computed result of
\(\alpha \in \mathbf{R}\) in floating point arithmetic by
\(\widehat{\alpha}\) or \(\fl{\alpha}\) and \(\mathbf{F}\) denotes the set of
all floating point numbers (see \cite{Higham2002} for more details).
Following \cite{Higham2002}, we will use the following classic properties in
error analysis.

@@

Now let us introduce some results concerning error-free transformation (EFT).
For a pair floating point numbers \(a, b \in \mathbf{F}\), there exists
\(y \in \mathbf{F}\) satisfying \(a \circ b = x + y\) where
\(x = \fl{a \circ b}\) and \(\circ \in \left\{+, -, \times\right\}\).
The transformation \((a, b) \longrightarrow (x, y)\) is regarded as an
error-free transformation. The error-free transformation algorithms of the
sum and product of two floating point numbers used later in this paper are
the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
algorithm by Dekker \cite{Dekker1971}, respectively. The following theorem
exhibits the important properties of \texttt{TwoSum} and \texttt{TwoProd}
algorithms (see \cite{Ogita2005}).

\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
\texttt{TwoSum} and \texttt{TwoProd} satisfy
\begin{itemize}
  \item \(\left[S, \sigma\right] = \mathtt{TwoSum}(a, b)\), \(x = \fl{a + b}\),
      \(S + \sigma = a + b\), \(\sigma \leq \mach \left|S\right|\),
      \(\sigma \leq \mach \left|a + b\right|\)
  \item \(\left[P, \pi\right] = \mathtt{TwoProd}(a, b)\),
      \(P = \fl{a \times b}\), \(P + \pi = a \times b\),
      \(\pi \leq \mach \left|P\right|\),
      \(\pi \leq \mach \left|a \times b\right|\).
\end{itemize}
The letters \(\sigma\) and \(\pi\) are used to indicate that the
errors came from sum and product, respectively. See the appendix
(Section~\ref{sec:appendix-algo}) for implementation details.
\end{theorem}

@@

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}(s)}{\left|p(s)\right|},
\end{equation}
where \(B_{j, n}(s) \geq 0\) and
\(\widetilde{p}(s) \coloneqq \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s)\).

@@

\noindent With this EFT in place\footnote{And the related
\texttt{LocalError} in Algorithm~\ref{alg:local-error}}, we can
perform \((K - 1)\) error filtrations:

@@

To enable a bound on the \(K\)-order error \(\db{K}^{(0)}_0\), it's necessary
to understand the difference between the exact local errors \(\ell_{F, j}\)
and the computed equivalents \(\widehat{\ell}_{F, j}\). To do this, we define
\begin{equation}
\widetilde{\ell}_{F, j} \coloneqq \left|e_1\right| +
 + \cdots + \left|e_{5F - 2}\right| + \left|\rho \cdot
\cdb{F - 1}_j^{(k + 1)}\right|.
\end{equation}

@@

\begin{proof}
Since
\begin{equation}
\mathtt{CompDeCasteljau}(p, s, K) = \widehat{b}_0^{(0)} \oplus
\cdb{1}_0^{(0)} \oplus \cdots \oplus \cdb{K - 1}_0^{(0)},
\end{equation}
applying Theorem~\ref{thm:sum-k} tells us that
\begin{equation}\label{sum-k-applied}
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right|.
\end{equation}
Since
\begin{equation}
p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
= \cdots
= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
+ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
\end{equation}
we have
\begin{gather}
\left|\sum_{F = 0}^{K - 1} \cdb{F}_0^{(0)}\right|
\leq \left|p(s)\right| + \left|\db{K}_0^{(0)}\right| \quad \text{and} \\
\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \leq
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\left|\db{K}_0^{(0)}\right| \label{triangle-ps}.
\end{gather}
Due to Lemma~\ref{lemma:k-order}, \(\db{F}_0^{(0)} =
\bigO{\mach^F} \widetilde{p}(s)\), hence
\begin{align}
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| &\leq
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right| &\leq
\gamma_{2n - 2}^K \left|\widehat{b}_0^{(0)}\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
&\leq
\gamma_{2n - 2}^K \left[\left|p(s)\right| +
  \bigO{\mach} \widetilde{p}(s)\right] +
\bigO{\mach^{K + 1}} \widetilde{p}(s).
\end{align}
Combining this with \eqref{sum-k-applied} and \eqref{triangle-ps}, we
see
\begin{align}
& \left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \\
\leq &
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\left|\db{K}_0^{(0)}\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
\leq &
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\left[\left(3^{K} \binom{n}{K} + \bigO{n^{K - 1}}\right) \mach^K +
\bigO{\mach^{K + 1}} \right]
\widetilde{p}(s).
\end{align}
Dividing this by \(\left|p(s)\right|\), we have our result.
\end{proof}

@@

Figure~\ref{fig:compensated-k} shows the relative forward errors compared
against the condition number. To compute relative errors, each input and
coefficient is converted to a fraction (i.e. infinite precision) and then
compared to the corresponding computed values. Similar tools are used to
\textbf{exactly} compute the condition number, though here we can rely
on the fact that \(\widetilde{p}(s) = (s - 1)
\left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and
condition numbers are computed as fractions, they are rounded to the
nearest IEEE-754 double precision value. As in \cite{Jiang2010}, we use
values \(\left\{\frac{3}{4} - (1.3)^j\right\}_{j=-5}^{-90}\)\footnote{As with
\(0.1\), it's worth noting that \((1.3)^j\) can't be represented exactly in
IEEE-754 double precision. However, this geometric series still serves a
useful purpose since it continues to raise \(\cond{p, s}\) as \(j\) decreases
away from \(0\) and because it results in ``random'' changes in the bits of
\(0.75\) that are impacted by subtracting \((1.3)^j\).}. The curves for
\texttt{DeCasteljau} and \texttt{CompDeCasteljau} trace the same paths seen
in \cite{Jiang2010}. In particular, \texttt{CompDeCasteljau} has a relative
error that is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
\(1 / \mach\), at which point the relative error increases linearly with
the condition number until it becomes \(\bigO{1}\) when
\(\cond{p, s}\) reaches \(1 / \mach^2\).
Similarly, the relative error in \texttt{CompDeCasteljau3}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 3\))
is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
\(1 / \mach^2\) at which point the relative error increases linearly
to \(\bigO{1}\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
and the relative error in \texttt{CompDeCasteljau4}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 4\))
is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
\(1 / \mach^3\) at which point the relative error increases linearly
to \(\bigO{1}\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).

@@

Find here concrete implementation details on the EFTs described
in Theorem~\ref{thm:eft}. They do not use branches, nor access to the
mantissa that can be time-consuming

@@

\noindent The following algorithms from \cite{Ogita2005} can be used as a
compensated method for computing a sum of \(n\) numbers. The first is a vector
transformation that is used as a helper:

@@

  \begin{algorithmic}
    \Function{\(\mathtt{res} = \mathtt{VecSum}\)}{$p$}
      \State \(n = \texttt{length}(p)\)
      \For{\(j = 2, \ldots, n\)}
        \State \(\left[p_i, p_{i - 1}\right] = \mathtt{TwoSum}\left(
            p_i, p_{i - 1}\right)\)
      \EndFor
    \EndFunction
  \end{algorithmic}
\end{breakablealgorithm}

@@

\ldots, e_{5F - 2}'\). In the same fashion as above, we track where
the current errors come from:

@@

For example, since we have \(r_1(k) = 3 \binom{k}{0}\) we'll have
\(q_1(k) = 3 \binom{k}{1}\). Once this is known
\begin{equation}
r_2(k) = 3 q_1(k - 1) + 5 r_1(k) = 3 \cdot 3 \binom{k - 1}{1} +
5 \cdot 3 \binom{k}{0} = 9 \binom{k}{1} + 6 \binom{k}{0}.
\end{equation}
If we write these polynomials in the ``falling factorial'' basis of
forward differences, then we can show that
\begin{equation}
r_F(k) = 3^F \binom{k}{F} + \cdots
\end{equation}
which will complete the proof. To see this, first note that for a polynomial
in this basis
\(f(k) = A \binom{k}{d} + B \binom{k}{d - 1} + C \binom{k}{d - 2} +
D \binom{k}{d - 3} + \cdots\) we have
\begin{align}
f(k + 1) &= A \binom{k}{d} + (A + B) \binom{k}{d - 1} +
  (B + C) \binom{k}{d - 2} + (C + D) \binom{k}{d - 3} + \cdots \\
f(k - 1) &= A \binom{k}{d} + (B - A) \binom{k}{d - 1} +
  (C - B + A) \binom{k}{d - 2} + (D - C + B - A) \binom{k}{d - 3} + \cdots
\end{align}
Using these, we can show that if
\(r_F(k) = \sum_{j = 0}^{F - 1} c_j \binom{k}{j}\) then
\begin{align}
q_F(k) &= c_{F - 1} \binom{k}{F} + \sum_{j = 1}^{F - 1}
  (c_j + c_{j - 1}) \binom{k}{j} \\
r_{F + 1}(k) &= 3 \left[-c_0 \binom{k}{0} +
  \sum_{j = 1}^F c_{j - 1} \binom{k}{j}\right] +
(5F) \sum_{j = 0}^{F - 1} c_j \binom{k}{j} =
3 c_{F - 1} \binom{k}{F} + \cdots
\end{align}
Under the inductive hypothesis \(c_{F - 1} = 3^F\) so that
the lead term in \(r_{F + 1}(k)\) is \(3 c_{F - 1} \binom{k}{F}
= 3^{F + 1} \binom{k}{F}\).

For the second inequality, we'll show that
\begin{equation}
\sum_{k = 0}^{n - 1} \gamma_{3k + 5F} L_{F, k} \leq
  \left[q_{F + 1}(n) \mach^{F + 1} +
  \bigO{\mach^{F + 2}}\right] \cdot \widetilde{p}(s)
\end{equation}
and then we'll have our result since we showed above that
\(q_{F + 1}(n) = 3^{F + 1} \binom{n}{F + 1} + \bigO{n^F}\). Since
\(\gamma_{3k + 5F} L_{F, k} \leq (3k + 5F) L_{F, k} \mach +
\bigO{\mach^{F + 1}} \widetilde{p}(s)\) it's enough to consider
\begin{equation}
\sum_{k = 0}^{n - 1} (3k + 5F) r_F(n - k) =
\sum_{k = 1}^n (3(n - k) + 5F) r_F(k).
\end{equation}
Since \(q_F(k) = q_F(k - 1) + r_F(k)\) and \(q_F(0) = 0\) we have
\(q_{F}(n) = \sum_{k = 1}^n r_{F}(k)\) thus
\begin{equation}
q_{F + 1}(n) = \sum_{k = 1}^n r_{F + 1}(k)
= \sum_{k = 1}^n 3 q_F(k - 1) + 5 F r_F(k)
= \sum_{k = 1}^n 3 \left[\sum_{j = 1}^{k - 1} r_F(j)\right] + 5 F r_F(k).
\end{equation}
Swapping the order of summation and grouping like terms, we have our
result.
\end{proof}

>>>>

As in~\eqref{de-casteljau-error}, the accuracy of the compensated
result~\eqref{de-casteljau-2-error} may be arbitrarily bad for ill-conditioned
polynomial evaluations. For example, as the condition number grows in
Figure~\ref{fig:jlcs-10}, some points have relative error exactly equal to
\(1\); this indicates that \(\mathtt{CompDeCasteljau}(p, s) = 0\), which is
a complete failure to evaluate the order of magnitude of \(p(s)\). For
root-finding problems \(\mathtt{CompDeCasteljau}(p, s) = 0\) when
\(p(s) \neq 0\) can cause premature convergence and incorrect results.
We describe how to defer rounding into progressively
smaller error groups and improve the accuracy of the computed result by a
factor of \(\mach\) for every error group added. So we derive
\texttt{CompDeCasteljauK}, a \(K\)-fold compensated de Casteljau algorithm
that satisfies the following a prior bound for any arbitrary integer \(K\):
\begin{equation}
  \frac{\left|p(s) - \mathtt{CompDeCasteljauK}(p, s, K)\right|}{
    \left|p(s)\right|} \leq \mach + \cond{p, s} \times
    \bigO{\mach^K}.
\end{equation}
This means that the computed value with \texttt{CompDeCasteljauK} is now
as accurate as the result of the de Casteljau algorithm performed in
\(K\) times the working precision with a final rounding back to the
working precision.

@@

We assume all floating point operations obey
\begin{equation}
  a \star b = \fl{a \circ b} = (a \circ b)(1 + \delta_1) =
  (a \circ b) / (1 + \delta_2)
\end{equation}
where \(\star \in \left\{\oplus, \ominus, \otimes, \oslash\right\}\), \(\circ
\in \left\{+, -, \times, \div\right\}\) and \(\left|\delta_1\right|,
\left|\delta_2\right| \leq \mach\). The symbol \(\mach\) is the unit round-off
and \(\star\) is a floating point operation, e.g.
\(a \oplus b = \fl{a + b}\). (For IEEE-754 floating point double precision,
\(\mach = 2^{-53}\).) We denote the computed result of
\(\alpha \in \mathbf{R}\) in floating point arithmetic by
\(\widehat{\alpha}\) or \(\fl{\alpha}\) and use \(\mathbf{F}\) as the set of
all floating point numbers (see \cite{Higham2002} for more details).
Following \cite{Higham2002}, we will use the following classic properties in
error analysis.

@@

An error-free transformation is a computational method where both
the computed result and the round-off error are returned. It
is considered ``free'' of error if the round-off can be represented
exactly as an element of \(\mathbf{F}\).
The error-free transformations used in this paper are
the \texttt{TwoSum} algorithm by Knuth \cite{Knuth1997} and \texttt{TwoProd}
algorithm by Dekker \cite{Dekker1971}, respectively.

\begin{theorem}[\cite{Ogita2005}]\label{thm:eft}
For \(a, b \in \mathbf{F}\) and \(P, \pi, S, \sigma \in \mathbf{F}\),
\texttt{TwoSum} and \texttt{TwoProd} satisfy
\begin{alignat}{4}
\left[S, \sigma\right] &= \mathtt{TwoSum}(a, b), & \, S &= \fl{a + b},
  S + \sigma &= a + b, \sigma &\leq \mach \left|S\right|,
  & \, \sigma &\leq \mach \left|a + b\right| \\
\left[P, \pi\right] &= \mathtt{TwoProd}(a, b),
  & \, P &= \fl{a \times b}, P + \pi &= a \times b,
  \pi &\leq \mach \left|P\right|,
  & \, \pi &\leq \mach \left|a \times b\right|.
\end{alignat}
The letters \(\sigma\) and \(\pi\) are used to indicate that the
errors came from sum and product, respectively. See the appendix
(Section~\ref{sec:appendix-algo}) for implementation details.
\end{theorem}

@@

The relative condition number of the evaluation of \(p(s) = \sum_{j = 0}^n
b_j B_{j, n}(s)\) in Bernstein form used in this paper is (see
\cite{Mainar1999}, \cite{Farouki1987}):
\begin{equation}
\cond{p, s} = \frac{\widetilde{p}(s)}{\left|p(s)\right|},
\end{equation}
where \(B_{j, n}(s) = \binom{n}{j} (1 - s)^{n - j} s^j \geq 0\) and
\(\widetilde{p}(s) \coloneqq \sum_{j = 0}^n \left|b_j\right| B_{j, n}(s)\).

@@

\noindent With this EFT in place\footnote{And the related
\texttt{LocalError} in Algorithm~\ref{alg:local-error}}, we can
perform \((K - 1)\) error filtrations. Once we've computed the \(K\) stages
of global errors, they can be combined with
\texttt{SumK} (Algorithm~\ref{alg:sum-k}) to produce a sum that is as
accurate as if computed in \(K\) times the working precision.

@@

To enable a bound on the \(K\)-order error \(\db{K}^{(0)}_0\), it's necessary
to understand the difference between the exact local errors \(\ell_{F, j}\)
and the computed equivalents \(\widehat{\ell}_{F, j}\). To do this, we define
\begin{equation}
\widetilde{\ell}_{F, j} \coloneqq \left|e_1\right| +
\cdots + \left|e_{5F - 2}\right| + \left|\rho \cdot
\cdb{F - 1}_j^{(k + 1)}\right|.
\end{equation}

@@

\begin{proof}
Since
\begin{equation}
\mathtt{CompDeCasteljau}(p, s, K) = \mathtt{SumK}\left(\left[
  \widehat{b}_0^{(0)}, \ldots, \cdb{K - 1}_0^{(0)}\right], K\right),
\end{equation}
applying Theorem~\ref{thm:sum-k} tells us that
\begin{equation}\label{sum-k-applied}
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| \leq
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right|.
\end{equation}
Since
\begin{equation}
p(s) = b_0^{(0)} = \widehat{b}_0^{(0)} + \db{1}_0^{(0)}
= \cdots
= \widehat{b}_0^{(0)} + \cdb{1}_0^{(0)} + \cdots
+ \cdb{K - 1}_0^{(0)} + \db{K}_0^{(0)}
\end{equation}
we have
\begin{gather}
\left|\sum_{F = 0}^{K - 1} \cdb{F}_0^{(0)}\right|
\leq \left|p(s)\right| + \left|\db{K}_0^{(0)}\right| \quad \text{and} \\
\left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \leq
\left|\mathtt{CompDeCasteljau}(p, s, K) - \sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| +
\left|\db{K}_0^{(0)}\right| \label{triangle-ps}.
\end{gather}
Due to Lemma~\ref{lemma:k-order}, \(\db{F}_0^{(0)} =
\bigO{\mach^F} \widetilde{p}(s)\), hence
\begin{align}
\left(\mach + 3 \gamma_{n - 1}^2\right) \left|\sum_{F = 0}^{K - 1}
\cdb{F}_0^{(0)}\right| &\leq
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
\gamma_{2n - 2}^K \sum_{F = 0}^{K - 1} \left|\cdb{F}_0^{(0)}\right| &\leq
\gamma_{2n - 2}^K \left|\widehat{b}_0^{(0)}\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
&\leq
\gamma_{2n - 2}^K \left[\left|p(s)\right| +
  \bigO{\mach} \widetilde{p}(s)\right] +
\bigO{\mach^{K + 1}} \widetilde{p}(s).
\end{align}
Combining this with \eqref{sum-k-applied} and \eqref{triangle-ps}, we
see
\begin{align}
& \left|\mathtt{CompDeCasteljau}(p, s, K) - p(s)\right| \\
\leq &
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\left|\db{K}_0^{(0)}\right| +
\bigO{\mach^{K + 1}} \widetilde{p}(s) \\
\leq &
\left[\mach + \bigO{\mach^2}\right] \left|p(s)\right| +
\left[\left(3^{K} \binom{n}{K} + \bigO{n^{K - 1}}\right) \mach^K +
\bigO{\mach^{K + 1}} \right]
\widetilde{p}(s).
\end{align}
Dividing this by \(\left|p(s)\right|\), we have our result.
\end{proof}

@@

Figure~\ref{fig:compensated-k} shows the relative forward errors compared
against the condition number. To compute relative errors, each input and
coefficient is converted to a fraction (i.e. infinite precision) and
\(p(s)\) is computed exactly as a fraction, then
compared to the corresponding computed values. Similar tools are used to
\textbf{exactly} compute the condition number, though here we can rely
on the fact that \(\widetilde{p}(s) = (s - 1)
\left(\frac{s}{2} - \frac{3}{4}\right)^7\). Once the relative errors and
condition numbers are computed as fractions, they are rounded to the
nearest IEEE-754 double precision value. As in \cite{Jiang2010}, we use
values \(\left\{\frac{3}{4} - (1.3)^j\right\}_{j=-5}^{-90}\)\footnote{As with
\(0.1\), it's worth noting that \((1.3)^j\) can't be represented exactly in
IEEE-754 double precision. However, this geometric series still serves a
useful purpose since it continues to raise \(\cond{p, s}\) as \(j\) decreases
away from \(0\) and because it results in ``random'' changes in the bits of
\(0.75\) that are impacted by subtracting \((1.3)^j\).}. The curves for
\texttt{DeCasteljau} and \texttt{CompDeCasteljau} trace the same paths seen
in \cite{Jiang2010}. In particular, \texttt{CompDeCasteljau} has a relative
error that is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
\(1 / \mach\), at which point the relative error increases linearly with
the condition number until it becomes \(\bigO{1}\) when
\(\cond{p, s}\) reaches \(1 / \mach^2\).
Similarly, the relative error in \texttt{CompDeCasteljau3}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 3\))
is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
\(1 / \mach^2\) at which point the relative error increases linearly
to \(\bigO{1}\) when \(\cond{p, s}\) reaches \(1 / \mach^3\)
and the relative error in \texttt{CompDeCasteljau4}
(Algorithm~\ref{alg:k-comp-de-casteljau} with \(K = 4\))
is \(\bigO{\mach}\) until \(\cond{p, s}\) reaches
\(1 / \mach^3\) at which point the relative error increases linearly
to \(\bigO{1}\) when \(\cond{p, s}\) reaches \(1 / \mach^4\).

@@

Find here concrete implementation details on the EFTs described
in Theorem~\ref{thm:eft}. They do not use branches, nor access to the
mantissa that can be time-consuming.

@@

\noindent The following algorithms from \cite{Ogita2005} can be used as a
compensated method for computing a sum of numbers. The first is a vector
transformation that is used as a helper:

@@

  \begin{algorithmic}
    \Function{\(\mathtt{res} = \mathtt{VecSum}\)}{$p$}
      \State \(n = \texttt{length}(p)\)
      \For{\(j = 2, \ldots, n\)}
        \State \(\left[p_j, p_{j - 1}\right] = \mathtt{TwoSum}\left(
            p_j, p_{j - 1}\right)\)
      \EndFor
    \EndFunction
  \end{algorithmic}
\end{breakablealgorithm}

@@

\ldots, e_{5F - 2}'\) that show up in \(\widetilde{\ell}_F\). In the same
fashion as above, we track where the current errors come from:

@@

For example, since we have \(r_1(k) = 3 \binom{k}{0}\) we'll have
\(q_1(k) = 3 \binom{k}{1}\). Once this is known
\begin{equation}
r_2(k) = 3 q_1(k - 1) + 5 r_1(k) = 3 \cdot 3 \binom{k - 1}{1} +
5 \cdot 3 \binom{k}{0} = 9 \binom{k}{1} + 6 \binom{k}{0}.
\end{equation}
If we write these polynomials in the ``falling factorial'' basis of
forward differences, then we can show that
\begin{equation}
r_F(k) = 3^F \binom{k}{F} + \cdots
\end{equation}
which will complete the proof of the first inequality. To see this, first
note that for a polynomial in this basis
\(f(k) = A \binom{k}{d} + B \binom{k}{d - 1} + C \binom{k}{d - 2} +
D \binom{k}{d - 3} + \cdots\) we have
\begin{align}
f(k + 1) &= A \binom{k}{d} + (A + B) \binom{k}{d - 1} +
  (B + C) \binom{k}{d - 2} + (C + D) \binom{k}{d - 3} + \cdots \\
f(k - 1) &= A \binom{k}{d} + (B - A) \binom{k}{d - 1} +
  (C - B + A) \binom{k}{d - 2} + (D - C + B - A) \binom{k}{d - 3} + \cdots
\end{align}
Using these, we can show that if
\(r_F(k) = \sum_{j = 0}^{F - 1} c_j \binom{k}{j}\) then
\begin{align}
q_F(k) &= c_{F - 1} \binom{k}{F} + \sum_{j = 1}^{F - 1}
  (c_j + c_{j - 1}) \binom{k}{j} \\
r_{F + 1}(k) &= 3 \left[-c_0 \binom{k}{0} +
  \sum_{j = 1}^F c_{j - 1} \binom{k}{j}\right] +
5F \left[\sum_{j = 0}^{F - 1} c_j \binom{k}{j}\right] =
3 c_{F - 1} \binom{k}{F} + \cdots
\end{align}
Under the inductive hypothesis \(c_{F - 1} = 3^F\) so that
the lead term in \(r_{F + 1}(k)\) is \(3 c_{F - 1} \binom{k}{F}
= 3^{F + 1} \binom{k}{F}\).

For the second inequality, we'll show that
\begin{equation}
\sum_{k = 0}^{n - 1} \gamma_{3k + 5F} L_{F, k} \leq
  \left[q_{F + 1}(n) \mach^{F + 1} +
  \bigO{\mach^{F + 2}}\right] \cdot \widetilde{p}(s)
\end{equation}
and then we'll have our result since we showed above that
\(q_{F + 1}(n) = 3^{F + 1} \binom{n}{F + 1} + \bigO{n^F}\). Since
\(\gamma_{3k + 5F} L_{F, k} \leq (3k + 5F) L_{F, k} \mach +
\bigO{\mach^{F + 2}} \widetilde{p}(s)\) it's enough to consider
\begin{equation}
\sum_{k = 0}^{n - 1} (3k + 5F) r_F(n - k) =
\sum_{k = 1}^n (3(n - k) + 5F) r_F(k).
\end{equation}
Since \(q_F(k) = q_F(k - 1) + r_F(k)\) and \(q_F(0) = 0\) we have
\(q_{F}(n) = \sum_{k = 1}^n r_{F}(k)\) thus
\begin{equation}
q_{F + 1}(n) = \sum_{k = 1}^n r_{F + 1}(k)
= \sum_{k = 1}^n 3 q_F(k - 1) + 5 F r_F(k)
= \sum_{k = 1}^n 3 \left[\sum_{j = 1}^{k - 1} r_F(j)\right] + 5 F r_F(k).
\end{equation}
Swapping the order of summation and grouping like terms, we have our
result.
\end{proof}
